---
title: "bis557-hw1"
author: "Name: Brian Deng"
output: rmarkdown::pdf_document
# output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bis557-hw1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bis557)
```

# Name: Brian Deng (BIS557 HW1)

The R package {bis557} was built to store all the functions and data used for 
this class, which focuses on creating statistical models and algorithms. 

- `linear_model()` - create a linear model, similar to R's `lm()` function
- `grad_descent()` - the gradient descent algorithm used in ML

## Dataset

Below is an example dataset, `lm_patho` from the {bis557} package, that can be 
used for linear regression.

```{r dataset lm_patho}
library(bis557)
data(lm_patho)
head(lm_patho)
```

## Example: Linear Model

This is a basic example which shows you to create a linear model, for example, 
doing regression analysis. 

$$
y \sim \beta_1 x_1 + \beta_2 x_2
$$

<br><br>

```{r function linear_model}
library(bis557)
data(lm_patho)
fit_linear_model <- linear_model(y ~., lm_patho)
```

## Example: Gradient Descent

This is a basic example which shows you how to solve a common problem: Use an 
optimization algorithm, such as gradient descent, to find the coefficients of 
simple linear regression. 

<br><br>

```{r function grad_descent}
library(bis557)
data(lm_patho)
gd_patho <- grad_descent(X = lm_patho[,-1], y = lm_patho[,1],
                         b_0 = rep(1e-16, ncol(lm_patho)), learn_rate = 1.3e-16,
                         max_iter = 1e5)
gd_patho
```
